---
title: "Covid19 vaccine study Fall2021"
author: "Christina Jilavdaryan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="fall21.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.


# Methods

Firstly we have to isolate the population categories 



# Males

```{r}
maledata <- filter(FinalData, sex=="guy") 
maledata$sex <- as.factor(maledata$sex)
```

## Bar


```{r}
Males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


Based on the data from the bar chart it seems that out of about 340 males who were infected with the covid virus, about 57% of the males who had covid were given the treatment, while around 43% of the males were given the placebo instead. 



## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```
Based on this table, it appears that around 340 males were infected with the covid viru out of the close to 10,400 total males in this study. This table shows that 3.72% of males who recieved the treatment had gotten covid, while 2.55% of males who got the placebo instead got covid.  



# Females








# Drug Users



# LGBTQ





# Conclusions and recommendations

