---
title: "Covid19 vaccine study Fall2021"
author: "Christina Jilavdaryan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="fall21.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex =="gal")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

To start this study we have to isolate the subpopulations into Males, Females, LGBT, and Drug users and look into the effectiveness of the studies treatment for each population

# Methods

$H_0$: The vaccine made no difference in whether or not males got covid

$H_A$: The vaccine does make a difference in whether or not males got covid




# Males
Firstly we have to isolate the population categories 

```{r}
maledata <- filter(FinalData, sex=="guy") 
maledata$sex <- as.factor(maledata$sex)
```

## Bar


```{r}
Males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


Based on the data from the bar chart it seems that out of about 340 males who were infected with the covid virus, about 57% of the males who had covid were given the treatment, while around 43% of the males were given the placebo instead. 



## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```
Based on this table, it appears that around 337 males were infected with the covid virus out of the close to 10,787 total males in this study. This table shows that 3.72% of males who recieved the treatment had gotten covid, while 2.55% of males who got the placebo instead got covid. This table shows that males who had gotten the vaccine were about 50% more likely to contract covid, which suggests that this treatment is not effective for the prevention of covid in males.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The Chi squared table shows that the P-value of 0.0006181. This value is so small so it is unlikely that the results are by chance.

Therefore we reject the null hypothesis because  males who got the vaccine were more likely to get covid so there was a difference.

The odds ratio from the Fisher exact test shows 0.84:1, when taking the odds ratio and turning it into a percentage of effectiveness using the following formula: 0/(1+0)*100 you get 40% effectiveness. 

Ultimately the vaccine is around 40% effective for males which makes the treatment in this study ineffective for preventing covid in males. 

# Methods

$H_0$: The vaccine made no difference in whether or not females got covid

$H_A$: The vaccine does make a difference in whether or not females got covid

# Females




```{r}
femaledata <- filter(FinalData, sex=="gal") 
femaledata$sex <- as.factor(femaledata$sex)
```


## Bar


```{r}
Females <- femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


This bar chart shows that out of approximately 607 females who were infected with covid, 32% received the treatment, while 66% were given the placebo.


## Numerical Summary

```{r}
table2 <- xtabs(~treatment + infected, data=femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```


Based on this table, it appears that around 607 females were infected with the covid virus out of the close to 9,213 total females in this study. This table shows that 4.44% of females who recieved the treatment had gotten covid, while 8.67% of females who got the placebo instead got covid. This table shows that females who had gotten the vaccine were about 67% less likely to contract covid, which suggests that this treatment is effective for the prevention of covid in females.

## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The p-value is small at 0.00000000000000002 that it is highly unlikely that the result was by chance. 

Therefore we reject the null hypothesis because  females who got the vaccine were less likely to get covid so there was a difference.

Based on the Fisher exact test ratio of 2.05:1, then putting it into the formula of 0/(1+0)*100 you get 67% effectiveness. The vaccine is 67% effective for females, which means the treatment in this study is effective for prevention of covid in females.


# Drug Users


# Methods

$H_0$: The vaccine made no difference in whether or not drug users got covid

$H_A$: The vaccine does make a difference in whether or not drug users got covid


```{r}
Druggiedata <- filter(FinalData, DrugUser=="yes") 
Druggiedata$DrugUser <- as.factor(Druggiedata$DrugUser)
```




## Bar


```{r}
Druggie<- Druggiedata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggie,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggie,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```



This bar chart shows that out of approximately 117 drug users who were infected with covid, 33.3% received the treatment, while 66.6% were given the placebo.

## Numerical Summary

```{r}
table3 <- xtabs(~treatment + infected, data=Druggiedata)
rowPerc(table3)
colPerc(table3)
print(table3)
```



Based on this table, it appears that around 117 Drug Users were infected with the covid virus out of the close to 2,283 total drug users in this study. This table shows that 33.3% of drug users who recieved the treatment had gotten covid, while 66.6% of drug users who got the placebo instead got covid. This table shows that drug users who had gotten the vaccine were about 60% less likely to contract covid, which suggests that this treatment is effective for prevention of covid in drug users.

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```
The vaccine is 60% effective for the drug user population


The p-value is small at 0.002273 that it is highly unlikely that the result was by chance. 

Therefore we reject the null hypothesis because drug users who got the vaccine were more less likely to get covid so there was a difference.

Based on the Fisher exact test ratio of 1.83:1, then putting it into the formula of 0/(1+0)*100 you get 60% effectiveness. The vaccine is 60% effective for drug users, which means the treatment in this study is effective for prevention of covid in drug users.



# LGBTQ



# Methods

$H_0$: The vaccine made no difference in whether or not LGBTQ got covid

$H_A$: The vaccine does make a difference in whether or LGBTQ got covid


```{r}
LGBTQdata <- filter(FinalData, LGBTQ=="gay") 
LGBTQdata$LGBTQ <- as.factor(LGBTQdata$LGBTQ)
```




## Bar


```{r}
LGBTQ<- LGBTQdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart shows that out of approximately 116 drug users who were infected with covid, 70% received the treatment, while 29% were given the placebo.


## Numerical Summary

```{r}
table4 <- xtabs(~treatment + infected, data=LGBTQdata)
rowPerc(table4)
colPerc(table4)
print(table4)
```
116  out of 2042 LGBTQ



Based on this table, it appears that around 116 LGBTQ were infected with the covid virus out of the close to 2,042 total LGBTQ in this study. This table shows that 70% of LGBTQ who recieved the treatment had gotten covid, while 29% of LGBTQ who got the placebo instead got covid. This table shows that LGBTQ who had gotten the vaccine were about 70% more ikely to contract covid, which suggests that this treatment is ineffective for prevention of covid in LGBTQ.


## Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```


The p-value is small at 0.0000057 that it is highly unlikely that the result was by chance. 

Therefore we reject the null hypothesis because  LGBTQ who got the vaccine were more likely to get covid so there was a difference.

Based on the Fisher exact test ratio of 0.40:1, then putting it into the formula of 0/(1+0)*100 you get 29% effectiveness. The vaccine is 29% effective for LGBTQ, which means the treatment in this study is ineffective for prevention of covid in LGBTQ.



# Conclusions and recommendations

When looking at the  four populations of male, female, drug user, and LGBTQ, we found that two populations had the highest rate of effectiveness when taking the vaccine for the prevention of covid which were females at 67% effectiveness and drug users at 60% effectiveness. We also found that two populations had the lowest rate of effectiveness in the prevention of covid after taking the vaccine which were males at 40% effectiveness and LGBTQ at 29% effectiveness. For males and LGBTQ the vaccine was so inneffective it actually increased the chances of getting covid.

